---
title: "Practical Machine Learning Course Project"
author: "Tze Yen Wong"
date: "09/10/2020"
output: html_document
---

## Overview
This report looks at devices such as Jawbone Up, Nike FuelBand, and Fitbit to predict the manner in which user did the exercise. A group of enthusiasts take measurements about themselves regularly to improve their health, to find patterns in their behaviour and to quantify how much of a particular activity they do. In this report, we use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants which they perform barbell lifts correctly and incorrectly in 5 different ways. To predict, the "classe" variable in the training set should be use along with other variables.


## Dataset
The data from the report come from:  
http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.

The training data:  
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data:  
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


## Loading packages
Loading packages needed for the analysis:
```{r, echo=TRUE}
library(caret)
library(rattle)
library(rpart)
set.seed(807)
```

## Loading and Cleaning Data
Loading data:
```{r, echo=TRUE}
train_data <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv')
validation <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv')
```

We partition the training data into 2 to create training (70%) and testing (30%) part.
```{r, echo=TRUE}
inTrain <- createDataPartition(train_data$classe, p=0.7, list=FALSE)
training <- train_data[inTrain,]
testing <- train_data[-inTrain,]
dim(training)
dim(testing)
```

Since the data may contain a lot of NA values and variables, we need to remove them before running analysis.
```{r, echo=TRUE}
# remove nearly zero variance variables
nearzero <- nearZeroVar(training)
training <- training[, -nearzero]
testing <- testing[,-nearzero]
dim(training)
dim(testing)

# remove mostly NA variables
var_NA <- sapply(training, function(x) mean(is.na(x)))
training <- training[, var_NA==FALSE]
testing <- testing[, var_NA==FALSE]
dim(training)
dim(testing)

# remove identification variables
training <- training[, -(1:5)]
testing <- testing[, -(1:5)]
dim(training)
dim(testing)
```

## Prediction Model Building 
For this project, we use 3 different model algorithm to predict and see which gives the highest accuracy. These include:   
1. Decision tree  
2. Generalized boosted model  
3. Random Forest  

### 1. Decision tree
```{r, echo=TRUE}
# Decision tree model fit
modFit_DT <- train(classe ~., method="rpart", data=training)
fancyRpartPlot(modFit_DT$finalModel)
```

### 2. Generalized boosted model
```{r, echo=TRUE}
# GBM model fit
control <- trainControl(method="repeatedcv", number=5, repeats=1)
modFit_GBM <- train(classe ~., method="gbm",data=training, trControl=control, verbose=FALSE)
modFit_GBM$finalModel
```

### 3. Random Forest 
```{r, echo=TRUE}
# Random forest model fit
modFit_RF <- train(classe ~., method="rf",data=training, trControl=trainControl(method="cv", 5), ntree=100)
modFit_RF$finalModel
```


## Model Assessment (Sample error)
```{r, echo=TRUE}
# Prediction on test data (Decision Tree)
predict_DT <- predict(modFit_DT, newdata=testing)
confm_DT <- confusionMatrix(factor(predict_DT), factor(testing$classe))
confm_DT

# Prediction on test data (GBM)
predict_GBM <- predict(modFit_GBM, newdata=testing)
confm_GBM <- confusionMatrix(factor(predict_GBM), factor(testing$classe))
confm_GBM

# Prediction on test data (Random Forest)
predict_RF <- predict(modFit_RF, newdata=testing)
confm_RF <- confusionMatrix(factor(predict_RF), factor(testing$classe))
confm_RF
```

## Applying the Selected Model to the Test Data
From the analysis above, we see that the accuracy of the 3 regression modeling methods are:  
  
1. Decision Tree: 0.4953  
2. GBM: 0.9842  
3. Random Forest: 0.9986  
  
Therefore, the random forest model will be applied to predict the 20 
quiz results. 
```{r, echo=TRUE}
predict_validation <- predict(modFit_RF, newdata=validation)
predict_validation
```

